Wrinkling of a confined porous layer

### 2014-05-30 poster_Okinawa.tex
Coupes, volume, aire des plis Altitude du fond d'une cloque (n'avance plus après contact avec le plafond)

\begin{textblock}{3}(0,8.5)
	\begin{tikzpicture}
	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png}\\
	};
	\begin{scope}[above left, white, fill=gray]
		\node at (m-1-1.south east) {\SI{15}{\minute}};
		\node at (m-2-1.south east) {\SI{22}{\minute}};
		\node at (m-3-1.south east) {\SI{1}{\hour}~26};
		\node at (m-4-1.south east) {\SI{1}{\hour}~30};
		\node at (m-5-1.south east) {\SI{1}{\hour}~31};
		\node at (m-6-1.south east) {\SI{1}{\hour}~35};
		\node at (m-7-1.south east) {\SI{2}{\hour}~25};
		\node at (m-8-1.south east) {\SI{2}{\hour}~40};
	\end{scope}
	\draw[<->, ultra thick] ($(m-8-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
	\end{tikzpicture}

@@

\begin{textblock}{3}(4,8.5)
\includegraphics[width=\textwidth]{volume_cloque_t116.jpg}

\begin{tikzpicture}[ultra thick]
\begin{axis}[
	name=a,
	width=\textwidth, height=0.25\textwidth, scale only axis,
	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
	axis lines=none, xtick=\empty,
	]
	\addplot+[Accent1]{cos(x)+1};
	\addplot+[Accent1]{cos(x)+1.5};
	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
	\node[above] at (axis cs:-90,1.5) {$P_2$};
	\node[below right] at (axis cs:-90,1.25) {$P_1$};
	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
\end{axis}
\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle (\textwidth,-1em);
\draw[<->] ($(a.south west)+(0,-1.1em)$) -- +(\textwidth,0) node[midway, below] {$\lambda$};
\end{tikzpicture}

\begin{itemize}
\item Synæresis then swelling
\item First blister after $\approx\SI{1}{\hour}$ of swelling
\item Independent blisters
\item $h,H,e \ll \lambda \ll L$
\item Once in contact with ceiling, $\lambda$ fixed
\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
\item Independent of gravity
\end{itemize}
\end{textblock}

\begin{textblock}{3}(8,8.5)
\Subhead{Ruck in a rug analogy}\\
\textit{Kolinski et al 2009, Vella et al 2009}\\
Bending energy $B$ vs vertical stress $\sigma$:
\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]

@@

\Subhead{Poroelasticity}\\
Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p$)
\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
\end{textblock}

@@

>>>>

\begin{textblock}{3}(0,8.5)
	\begin{tikzpicture}
	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t000.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t011.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t110.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t116.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t117.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t123.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t200.png}\\
	\includegraphics[width=\columnwidth, height=0.061\columnwidth]{coupe_cloque_t221.png}\\
	};
	\begin{scope}[above left, white, fill=gray, inner ysep=0]
		\node at (m-1-1.south east) {\SI{15}{\minute}};
		\node at (m-2-1.south east) {\SI{22}{\minute}};
		\node at (m-3-1.south east) {\SI{1}{\hour}~26};
		\node at (m-4-1.south east) {\SI{1}{\hour}~30};
		\node at (m-5-1.south east) {\SI{1}{\hour}~31};
		\node at (m-6-1.south east) {\SI{1}{\hour}~35};
		\node at (m-7-1.south east) {\SI{2}{\hour}~25};
		\node at (m-8-1.south east) {\SI{2}{\hour}~40};
	\end{scope}
	\draw[<->, ultra thick] ($(m-8-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
	\end{tikzpicture}

@@

\begin{textblock}{3}(4,8.5)

\begin{tikzpicture}
\begin{axis}[
	scale only axis,
	width=\textwidth-4em,
	height=0.3\textwidth,
	xmin=0, xmax=1000, xtick={0,250, 500, 750}, xlabel={position (\si{\micro\metre})},
	ymin=0, ymax=50, ylabel={altitude (\si{\micro\metre})},
	cycle list={
		{Accent2}, 
		{black!15!Accent2},
		{black!30!Accent2},
		{black!45!Accent2},
		{black!60!Accent2},
		{black!75!Accent2},
		{black}
	},
	no marks,
	]
	\foreach \x in {2,3,..., 8}
		\addplot table[y index=\x]{alts_bottom_cloque.txt};

	\draw[ultra thick, Accent1] (axis cs:580,0) -- (axis cs:580,50);
\end{axis}
\end{tikzpicture}
\includegraphics[width=\textwidth]{volume_cloque_t116.jpg}

%\Subhead{Phenomenology}
\begin{itemize}
\item Synæresis then swelling
\item First blister after $\approx\SI{1}{\hour}$ of swelling
\item Independent blisters
\item $h,H,e \ll \lambda \ll L$
\item Once in contact with ceiling, $\lambda$ fixed
\item Cascade buckling\linebreak \hfill\textit{Roman \& Pocheau 1999}
\item Independent of gravity
\end{itemize}
\end{textblock}

\begin{textblock}{3}(8,8.5)
\begin{tikzpicture}[ultra thick]
\begin{axis}[
	name=a,
	width=\textwidth, height=0.25\textwidth, scale only axis,
	domain=-180:180, no markers, ymin=0, ymax=4,xmin=-180,xmax=180,
	axis lines=none, xtick=\empty,
	]
	\addplot+[Accent1]{cos(x)+1};
	\addplot+[Accent1]{cos(x)+1.5};
	\draw[->, ultra thick] (axis cs:90,1.25) -- +(45:0.05\textwidth) node[pos=1, above] {$\vec{v}$};
	\node[above] at (axis cs:-90,1.5) {$P_2$};
	\node[below right] at (axis cs:-90,1.25) {$P_1$};
	\draw[<->, ultra thick] (axis cs:0,0) -- (axis cs:0,2) node[midway, right] {$A$};
\end{axis}
\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle (\textwidth,-1em);
\draw[<->] ($(a.south west)+(0,-1.1em)$) -- +(\textwidth,0) node[midway, below] {$\lambda$};
\end{tikzpicture}

\Subhead{Ruck in a rug analogy}\\
\textit{Kolinski et al 2009, Vella et al 2009}\\
Bending energy $B$ vs vertical stress $\sigma$:
\[ \lambda^* = \left(\frac{B}{\sigma}\right)^{1/3}\text{, eg. }\sigma =\rho g h \]

@@

\Subhead{Poroelasticity}\\
Here $\sigma$ is a Darcy pressure due to solvent flow through the gel (pore size $\ell_p \approx \SI{4}{\micro\metre}$)
\[\sigma = \Delta P \sim \eta h v/\ell_p^2   \]
\end{textblock}

@@

\begin{textblock}{3}(0,17)
	\begin{tikzpicture}
	\matrix[matrix of nodes, matrix anchor=north east, inner sep=0, row sep=0.01\textwidth]  (m){
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t000.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t016.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t032.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t038.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t040.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t043.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t055.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t070.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t123.png}\\
	\includegraphics[width=\columnwidth, height=0.052\columnwidth]{coupe_plis_t332.png}\\
	};
	\begin{scope}[above left, white, fill=gray, inner ysep=0]
		\node at (m-1-1.south east) {\SI{15}{\minute}};
		\node at (m-2-1.south east) {\SI{23}{\minute}};
		\node at (m-3-1.south east) {\SI{32}{\minute}};
		\node at (m-4-1.south east) {\SI{35}{\minute}};
		\node at (m-5-1.south east) {\SI{36}{\minute}};
		\node at (m-6-1.south east) {\SI{38}{\minute}};
		\node at (m-7-1.south east) {\SI{44}{\minute}};
		\node at (m-8-1.south east) {\SI{53}{\minute}};
		\node at (m-9-1.south east) {\SI{1}{\hour}~21};
		\node at (m-10-1.south east) {\SI{3}{\hour}~15};
	\end{scope}
	\draw[<->, ultra thick] ($(m-10-1.south west)+(0,-0.25em)$) -- +(0.786\textwidth,0) node[midway, below] {\SI{1}{\milli\metre}};
	\end{tikzpicture}

	\begin{tikzpicture}
	\begin{groupplot}[%
		group style={
			group name=g, group size=1 by 2,
			xticklabels at=edge bottom,
			vertical sep=0,
			},
		xmin=0, xmax=180, xtick={0,30,...,150},
		extra tick style={grid=major},%
		extra x tick labels={},%
		scale only axis,
		width=\textwidth-4em,
		height=0.3\textwidth,
		ylabel absolute, every axis y label/.append style={anchor=base, yshift=1.25em}
		]

	\nextgroupplot[ylabel={Volume (\%)}, ymin=20, ymax=100, ytick={40,60,80,100}, every axis y label/.append style={xshift=0.5em}]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{1}*100}]{relative_volume_excess_area_plis.txt};
	%\addplot+[only marks,Accent2, mark=+] table[y expr={\thisrowno{1}*100}]{volume_rel_half_cas8_toi.txt};

	\nextgroupplot[ylabel={Excess area (\%)}, ymin=0, ymax=6, ytick={0,2,4}, 
	xlabel={time (min)}, every axis y label/.append style={xshift=-1em}]
	\addplot+[no marks,Accent1] table[x expr={\thisrowno{0}+15}, y expr={\thisrowno{2}*100}]{relative_volume_excess_area_plis.txt};
	%\addplot+[only marks,Accent2, mark=+] table{excess_area_pc_half_cas8_toi.txt};
	\end{groupplot}
	\end{tikzpicture}
\end{textblock}

\begin{textblock}{7}(4,17)
\begin{tikzpicture}
\begin{axis}[
	name=a,
	width=\textwidth, height=0.25\textwidth, scale only axis,
	domain=0:360, no markers, ymin=-4, ymax=4,xmin=0,xmax=360,
	axis lines=none, xtick=\empty,
	]
	\fill[gray!20] (axis cs:85,4) rectangle (axis cs:95,-4) (axis cs:265,4) rectangle (axis cs:275,-4);
	\addplot+[Accent1]{sin(x)+0.5};
	\addplot+[Accent1]{sin(x)-0.5};
	\addplot+[dashed, black]{0.5};
	\addplot+[dashed, black]{-0.5};
	\draw[<->, help lines] (axis cs:180,0.5) -- (axis cs:180,4) node[midway, left] {$H$}; 
	\draw[<->, help lines] (axis cs:30,-0.5) -- (axis cs:30,-4) node[midway, left] {$H$};
	\node[below] at (axis cs:90,-1.5) (pb1) {$P_1$};
	\node[above] at (axis cs:90,1.5) (ph2) {$P_2$};
	\node[above] at (axis cs:270,1.5) (ph1) {$P_1$};
	\node[below] at (axis cs:270,-1.5) (pb2) {$P_2$};
	\draw[thick, ->] (ph2) -- (axis cs:90,0.5) node[midway, left] {Darcy} node[midway, right] {$v$};
	\draw[thick, ->] (pb2) -- (pb1) node[midway, above] {Poiseuille} node[midway, below] {$u \sim \frac{\lambda}{H} v$};
\end{axis}
\fill[pattern=north east lines,pattern color=Accent2] (a.south west) rectangle +(\textwidth,-1em) (a.north west) rectangle +(\textwidth,1em);
\end{tikzpicture}

\begin{itemize}
\item The same $\Delta P$ can be relaxed by a Poiseuille flow
\[\Delta P \sim 12\eta\frac{\lambda}{H^2} u \sim 12\eta\frac{\lambda^2}{H^3} v\]
\item For the same wavelength
\[\frac{v_\text{Darcy}}{v_\text{Poiseuille}} \sim \frac{\ell_p^2\lambda^2}{hH^3} \approx 30\]
\end{itemize}
